---
title: "FA1_Skewness_Analysis.Rmd"
author: "Arvie M. Sinocruz"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Introduction
This document calculates Pearson's skewness coefficient for examination subjects and visualizes student scores using stem-and-leaf plots and boxplots.

## Load Required Libraries
```{r}
library(readr)
library(dplyr)
```

## Read Dataset
```{r}
data <- read.csv("results.csv")
head(data)
```

## Task 1: Calculate Pearson's Skewness
```{r}
calculate_skewness <- function(x) {
  mean_x <- mean(x, na.rm = TRUE)
  median_x <- median(x, na.rm = TRUE)
  sd_x <- sd(x, na.rm = TRUE)
  
  skewness <- (3 * (mean_x - median_x)) / sd_x
  return(skewness)
}

# Apply function to numerical columns (excluding IDs/names if present)
skew_values <- sapply(data[, -1], calculate_skewness)  # Adjust column selection if necessary

# Display results
skew_values
```

## Task 2a: Stem-and-Leaf Display by Gender
```{r}
# Split data by gender
male_data <- data %>% filter(Gender == "Male")
female_data <- data %>% filter(Gender == "Female")

# Display stem-and-leaf plots
cat("Stem-and-leaf plot for Male Students:")
stem(male_data$Scores)  # Replace 'Scores' with the actual column name

cat("\nStem-and-leaf plot for Female Students:")
stem(female_data$Scores)
```

## Task 2b: Boxplot by Gender
```{r}
boxplot(Scores ~ Gender, data = data, 
        main = "Boxplot of Scores by Gender",
        xlab = "Gender",
        ylab = "Scores",
        col = c("lightblue", "pink"))

# Add mean points
means <- tapply(data$Scores, data$Gender, mean, na.rm = TRUE)
points(1:length(means), means, col="red", pch=19)
```

## Conclusion
- Skewness values indicate the distribution's symmetry.
- The stem-and-leaf plot preserves raw data, while boxplots show overall distribution trends.


