---
title: "Formative Assessment 5"
author: "Mercado, Consuelo"
date: "2025-03-01"
output: pdf_document
---

*github link*: 

# ITEM 6

An email message can travel through one of three server routes. The percentage of errors in each of the servers and the percentage of messages that travel through each route are given. We assume that the servers are independent.

```{r}
library(knitr)
data <- data.frame(
  servers = 1:3,
  percent_msg = c(40, 25, 35),
  percent_error = c(1, 2, 1.5)
)

kable(data, caption = "Item 6 Table")
```

```{r}
# GIVEN
p1 <- 0.40  # Server 1
p2 <- 0.25  # Server 2
p3 <- 0.35  # Server 3

ES1 <- 0.01  # error server 1
ES2 <- 0.02  # error server 2
ES3 <- 0.015 # error server 3
```

### a. What is the probability of receiving an email containing an error?

Using the Law of Total Probability:

$$
P(E) = P(E | S1) P(S1) + P(E | S2) P(S2) + P(E | S3) P(S3)
$$

```{r}
# Compute probability of an error
prob_error <- (ES1 * p1) + (ES2 * p2) + (ES3 * p3)
prob_error
```

### b. What is the probability that a message will arrive without error?

```{r}
# Compute probability of no error
prob_noerror <- 1 - prob_error
prob_noerror
```

### c. If a message arrives without error, what is the probability that it was sent through server 1?

by Bayes' Theorem:

$$
P(S1 | E^c) = \frac{P(E^c | S1) P(S1)}{P(E^c)}
$$

where $P(E^c | S1) = 1 - P(E | S1)$.

```{r}
probs1 <- 1 - ES1
probs1_noerror <- (probs1 * p1) / prob_noerror
probs1_noerror
```

### Conclusion

-   The probability of receiving an email containing an error is `r round(prob_error, 4)`, or `r round(prob_error * 100, 2)`%.
-   The probability of receiving an email without error is `r round(prob_noerror, 4)`, or `r round(prob_noerror * 100, 2)`%.
-   The probability that a message arrives without error is `r round(probs1_noerror, 4)`, or `r round(probs1_noerror * 100, 2)`%.

# ITEM 9

### Given Data

```{r}
# Define given probabilities for graphics package purchase
not_buying<- 0.20  # Probability of not buying
buying <- 0.70  # Probability of buying
undecided <- 0.10  # Probability of being undecided

# Probabilities of upgrading given decision
notbuying_up <- 0.10  # Given not buying, probability of upgrading
buying_up <- 0.40  # Given buying, probability of upgrading
undecided_up <- 0.20  # Given undecided, probability of upgrading
```

### A. Probability that a manager will not upgrade their computer hardware (P(¬G))

Using the Law of Total Probability:

```{r}
total_prob <- (notbuying_up * not_buying) + (buying_up * buying) + (undecided_up * undecided)
not_up_total <- 1 - total_prob  # P(¬G)

total_prob
not_up_total
```

### B. Posterior Probability: P(B \| G)

Using Bayes’ Theorem:

```{r}
prob_post <- (buying_up * buying) / total_prob
prob_post
```

### C. 

```{r}
# P(B | ¬G)
prob_gcb <- 1 - buying_up
prob_bgc <- (prob_gcb * buying) / not_up_total

# P(C | G)
prob_cg <- (undecided_up * undecided) / total_prob

# P(¬C | ¬G)
prob_cc <- 1 - undecided
prob_gcc <- 1 - undecided_up
prob_ccgc <- (prob_gcc * prob_cc) / not_up_total

prob_bgc
prob_cg
prob_ccgc
```

## Tree Diagram

```{r}
library(webshot2)


library(data.tree)

# Create the tree structure
root <- Node$new("Manager")

notbuying <- root$AddChild("Not Buying 20% ")
buying <- root$AddChild("Buying 70%")
undecided <- root$AddChild("Undecided 10% ")

buying$AddChild("interested in upgrading 40% ")

# Customize node appearance
SetNodeStyle(root, fontname = "Times", fontsize = 11, shape = "rectangle")
SetNodeStyle(buying, color = "lightgreen")
SetNodeStyle(not_buying, color = "red")
SetNodeStyle(undecided, color = "yellow")

# Customize edge styles
SetEdgeStyle(root, color = "black", penwidth = 2)
SetEdgeStyle(buying, color = "darkgreen", penwidth = 2, arrowhead = "vee")

# Plot the styled tree
plot(root)
```

# ITEM 13

A malicious spyware can infect a computer system through the Internet or through email. The spyware comes through the Internet 70% of the time and 30% of the time, it gets in through email. If it enters via the Internet, the anti-virus detector will detect it with probability 0.6, and via email, it is detected with probability 0.8.

(a) What is the probability that this spyware infects the system?
(b) If the spyware is detected, what is the probability that it came through the Internet?

```{r}
# Given probabilities
prob_spy_I <- 0.7   # Probability that spyware comes through Internet
prob_spy_E <- 0.3   # Probability that spyware comes through Email
prob_detect_I <- 0.6 # Probability that spyware is detected via Internet
prob_detect_E <- 0.8 # Probability that spyware is detected via Email
```

### a. What is the probability that this spyware infects the system?

```{r}
# Probability that spyware is NOT detected
not_detect_I <- 1 - prob_detect_I   # Probability that it is not detected via Internet
not_detect_E <- 1 - prob_detect_E  # Probability that it is not detected via Email

# Total probability of infection
prob_infect <- (prob_spy_I * not_detect_I) + (prob_spy_E * not_detect_E)
prob_infect
```

### b. If the spyware is detected, what is the probability that it came through the Internet?

```{r} 
#Total probability of detection 
prob_detect <- (prob_detect_I * prob_spy_I) + (prob_detect_E * prob_spy_E)

 #Bayes' Theorem: P(Internet | Detected)

internet_detect <- (prob_detect_I * prob_spy_I) / prob_detect
internet_detect

```