---
title: "Analysis of Scores"
author: "Arvie M. Sinocruz and Consuelo B. Mercado"
date: "2025-01-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Part  A: Stem-and-Leaf Display

In this section, we will generate the stem-and-leaf displays for male and female scores based on the data given in Exercise 1.1.

```{r stem_and_leaf}
#Female and Male Scores
male_scores <- c(48, 49, 49, 30, 30, 31, 32, 35, 37, 41, 86, 42, 51, 53, 56, 42, 44, 50, 51, 65, 67, 51, 56, 58, 64, 64, 75)
female_scores <- c(57, 59, 78, 79, 60, 65, 68, 71, 75, 48, 51, 55, 56, 41, 43, 44, 75, 78, 80, 81, 83, 83, 85)

#Stem-and-Leaf Display fo Males
cat("\nMale  Stem-and-Leaf  Display:\n")
stem(male_scores)

#Stem-and-Leaf Display fo Females
cat("\nFemale  Stem-and-Leaf  Display:\n")
stem(female_scores)
```

##Part  B: Box-Plot Construction

In this section, we will contruct the boxplot for male and female scores based on the data given in Exercise 1.1.

```{r box-plot for each gender}
#Boxplot Construction for Female Score
 boxplot(female_scores, main = "Boxplot for Female Scores", ylab = "Scores", col = c("pink"), names = c("Female"))

#Boxplot Construction for Male Score
boxplot(male_scores, main = "Boxplot for Male Scores", ylab = "Scores", col = c("lightblue"), names = c("Male"))
