---
title: "Formative Assessment 1"
author: "MERCADO, C & SINOCRUZ, A"
date: "2025-02-01"

output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

# Question 1

1\. Write the skewness program, and use it to calculate the skewness
coefficient of the four examination subjects in results.txt
(results.csv). What can you say about these data?

Pearson has given an approximate formula for the skewness that is easier
to calculate than the exact formula given in Equation 2.1.

$$\text{Skewness} \approx \frac{3(\text{Mean} - \text{Median})}{\text{Standard Deviation}}$$

Write a program to calculate this and apply it to the data in
results.txt (results.csv). Is it a reasonable approximation? 

## STEPS:

### Load the necessary libraries and data

```{r}
library(readr)
library(moments)
df <- read_csv("C:\\Users\\CONSUELO B. MERCADO\\OneDrive\\Documents\\r fas\\results.csv")

```

### We need to specify the file path for this. And then type

```{r}
library(moments)
```

### This is a required package for the skewness function that we will use for the exact skewness part.Next, we calculated the mean, median, and standard deviation excluding the gender column and the NA values. These are the needed values for Pearson's approximation

```{r}
sapply(df[2:5], mean, na.rm = TRUE)    

sapply(df[2:5], median, na.rm = TRUE) 

sapply(df[2:5], sd, na.rm = TRUE)

```

### From the moments package, we use the skewness function to calculate the exact skewness.

```{r}
numeric_columns <- c("arch1", "prog1", "arch2", "prog2") 

df_numeric <- df[, numeric_columns] 

exact_skewness <- sapply(df_numeric, skewness, na.rm = TRUE)
exact_skewness
```

### Next, let's compute for the Pearson Skewness.

```{r}
pearson_skewness <- sapply(df_numeric, function(x) {(3 * (mean(x, na.rm = TRUE) - 
median(x, na.rm = TRUE))) / sd(x, na.rm = TRUE)})
pearson_skewness
```

### To see the comparison of the two, let's combine them using data frame.

```{r}
skewness_results <- data.frame(Subject = numeric_columns, Exact_Skewness = 
exact_skewness, Pearson_Skewness = pearson_skewness)
skewness_results
```

### Analysis

According to the research (Green et al., 2023), skewness is used to
describe the lack of symmetries in a data distribution. In this problem,
we calculate the *Exact Skewness* and *Pearson Approximation* for the
four subjects in results.csv. The first two subjects, arch1 and prog1
showed a little difference in their exact and pearson values. However,
in arch2, the exact skewness is positive while the other is negative.
Also in the fourth subject, prog2, there is a big difference in their
skewness. This shows that Pearson Skewness is reasonable approximation
for data distribution but it still needs other statistical method just
like the Exact Skewness to check its accuracy.

### Reference:


Green, J. L., Manski, S. E., Hansen, T. A., & Broatch, J. E. (2023,
January 1). *Descriptive statistics* (R. J. Tierney, F. Rizvi, & K.
Ercikan, Eds.). ScienceDirect; Elsevier.
<https://www.sciencedirect.com/science/article/abs/pii/B9780128186305100831>
