---
title: "FA2"
author: "Mercado, C & Sinocruz, A"
date: "2025-02-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Use R to illustrate the probability of getting:

(a) a head is 0.5 if a fair coin is tossed repeatedly;
(b) a red card is 0.5 if cards are drawn repeatedly with replacement from an awell-shuffled deck;
(c) an even number is 0.5 if a fair die is rolled repeatedly.

## A. A head is 0.5 if a fair coin is tossed repeatedly.

*First, we need to simulate coin tosses.*

```{r}
set.seed(123)
n <- 10000  # for the number of trials
coin_tosses <- sample(c("H", "T"), size = n, replace = TRUE, prob = c(0.5, 0.5))
```

*Then, compute for the probability of getting a head.*

```{r}
p_head <- sum(coin_tosses == "H") / n
p_head
```

*So, the proportion is 0.4923 or approximately 0.5.*

## B. A red card is 0.5 if cards are drawn repeatedly with replacement from an awell-shuffled deck.

*First, let's simulate again. In a standard deck of cards, there are 26 black cards and 26 red cards.*

```{r}

deck <- c(rep("Red", 26), rep("Black", 26)) 
draws <- sample(deck, size = n, replace = TRUE)
```

*Then let's compute for the probability.*

```{r}
p_red <- sum(draws == "Red") / n
p_red
```

*So, the proportion is 0.4996 or approximately 0.5.*

## C. An even number is 0.5 if a fair die is rolled repeatedly.

*Like what we did in the first two, let's simulate a die roll*

```{r}
die_roll <- sample(1:6, size = n, replace = TRUE)
```

*Then let's calculate the probability.*

```{r}
p_even <- sum(die_roll %% 2 == 0) / n
p_even
```

*So, the proportion is 0.4969 or approximately 0.5.*

## Summary

```{r}
sumtbl <- data.frame(
  Event = c("Coin Toss (Head)", "Drawing a Red Card", "Rolling an Even Number"),
  Theoretical_Probability = c(0.5, 0.5, 0.5),
  Simulated_Probability = c(p_head, p_red, p_even)
)
sumtbl

```

*From there, we can see that all the simulated probability is very close to 0.5.*
